---
Project Title: "Predictive Model to Determine the Fitness Exercise Manner"
Author: "R Koder"
Date: "March 25, 2016"
Output Document Type: HTML
---

The objective of the project is to to predict the manner which the exercise is done
Data for the project is from the link: http://groupware.les.inf.puc-rio.br/har.

## 1. Data Retrieval

Retrieve the Training and Test data CSVs and download to the local directory.

```{r}

if (!file.exists("./pml-training.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
        destfile = "./pml-training.csv")
}
if (!file.exists("./pml-testing.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
        destfile = "./pml-testing.csv")
}

```

## 2. Exploration


The training set consists of 19622 observations of 160 variables.

```{r}

exp_training <- read.csv("./pml-training.csv")

dim(exp_training)
```

## 3. Data Pre Processing

Some columns of the raw data set have empty strings and those columns have to be deleted. The columns X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window are not related to outcome and hence can be deleted.

```{r}
library(caret)

set.seed(12345)
trainingdata <- read.csv("pml-training.csv", stringsAsFactors=FALSE)
trainingdata$classe <- as.factor(trainingdata$classe)
trainingdata <- trainingdata[,-nearZeroVar(trainingdata)]
trainingdata <- trainingdata[,-c(1,2,3,4,5,6,7)]

```

There are many NA values in the data set, so, KnnImpute method to impute those values.Features can be standardized and PCA can be used to reduce features.

```{r}
iTraining <- createDataPartition(y=trainingdata$classe, p=0.75, list=FALSE)
trainingdata <- trainingdata[iTraining,]
testingset <- trainingdata[-iTraining,]

preprocessObject <- preProcess(trainingdata[,-length(trainingdata)],method=c("center", "scale", "knnImpute", "pca"), thresh=0.9)
cleaned_data <- predict(preprocessObject,trainingdata[,-length(trainingdata)])
```

## 4. Prediction

Once the data set is cleansed, Knn method  can be used to build model.Testing data can be used to evaluate the performance of the model. 
The accuracy is 0.973. 

```{r}
model_ft <- train(trainingdata$classe ~.,data=cleaned_data, method="knn")
test <- predict(preprocessObject, testingset[,-length(testingset)])
confusionMatrix(testingset$classe, predict(model_ft,test))
```


Result as predicted from test data,
```{r}
testingset <- read.csv("pml-testing.csv", stringsAsFactors=FALSE)
testingset <- testingset[,names(testingset) %in% names(trainingdata)]

testdat <- predict(preprocessObject, testingset)
predict_result <- predict(model_ft, testdat)
```
